Scattering and Pair Processes in General Relativistic Ray Tracing for Neutrinos

### 2016-10-14 paper.tex
Add comprehensive todo notes to the text

Neutrino radiation obeys the relativistic Boltzmann Equation
\todo{explain limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2) = C[f],
\end{equation}
where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and depends locally on the distribution function
$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
positon $x^\alpha$, and momentum $p_\beta$ of the neutrinos in question,
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For clarity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

@@

Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
the direction of $\ell_\beta$. We choose $\lambda=0$ at $x^\alpha_e$
and $\lambda=\lambda_o$ at $x^\alpha_o$.
According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
parameter by $ds=p^t d\lambda$.
\todo{show}

\subsection{The Rendering Equation}
\label{ssec:rendering_eqn}

\subsection{Neutrinosurface approximation}
\label{ssec:neutrinosurface}

\subsection{Moments of the Distribution Function}
\label{ssec:moments}

\subsection{Numerical Implementation}
\label{ssec:numerical}

@@

\subsection{Dummy Tests}
Maybe we don't need to present these tests, but make sure to establish
confidence in the basics:
gravitational redshift, Doppler shift, thermodynamic equilibrium.

\subsection{Homogenous Absorbing Star}
Test described in~\cite[Sec.~3.2]{smit1997-two_moment}

\subsection{Homogenous Scattering Star}
Test presented in~\cite{humm1971-grey_transfer}
and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.

\subsection{Post-Bounce Neutron Core}
Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
and in \cite[App.~E.6]{fouc2015-m1_nsbh},
and in~\cite{abdi2012-monte_carlo}.

\subsection{Disk Comparison to M1}
Compute total and relative number and energy luminosities for a disk evolved
in one of~\cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
Expect agreement within factors of a few (?).

@@

\subsection{Emission Spectra}
\label{ssec:spectra}
The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
Something like this is presented in~\cite[Fig.~10-11]{pere2014-nu_wind};
compare.

\subsection{Closure Relation}
\label{ssec:closure}
This is a little tricky. Unless we can come up with a new improved analytic
closure relation and use ray tracing to evaluate its trustworthiness,
it's going to be hard to get anything useful from computing
$P^{\alpha\beta}$.

Maybe the variable Eddington closure is more suitable to this application.
If I understand it right, applying ray tracing to that method would require
computing $P^{\alpha\beta}$ on the fly throughout the spatial volume.
That would certainly be doable for a quasi-stationary disk configuration.

@@

\section{Source Terms}
\label{sec:source_terms}
Here we present the source terms comprising the right hand side of the
Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) in terms of the microphysical
processes of absorption/emission, elastic scattering, inellastic scattering,
and pair annihilation/production
(detailed in Tab.~\ref{tab:neutrino_processes}).

>>>>

Neutrino radiation obeys the relativistic Boltzmann Equation
\todo{point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2) = C[f],
\end{equation}
where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and depends locally on the distribution function
$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
positon $x^\alpha$, and momentum $p_\beta$ of the neutrinos in question,
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For clarity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

@@

Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
the direction of $\ell_\beta$. We choose $\lambda=\lambda_0$ at $x^\alpha_e$
and $\lambda=\lambda$ at $x^\alpha_o$, as in Fig.~\ref{fig:affine_param}.
\todo{choose a better notation}
According to Eqn.~\ref{eqn:geodesic_x} proper distance relates to the affine
parameter by $ds=p^t d\lambda$.
\todo{show explicitly}

\begin{figure}
  \includegraphics[width=\columnwidth]{affine_parameter_label}
  \caption{Affine parameterization of a trajectory.}
  \label{fig:affine_param}
\end{figure}

\subsection{The Rendering Equation}
\label{ssec:rendering_eqn}
\todo{type up derivation from notes}
We can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
\begin{equation}
  \label{eqn:rendering}
  f(\lambda) = f(\lambda_0) e^{-\tau(\lambda,\lambda_0)} +
  \int_{\lambda_0}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
\end{equation}
where the parameterization conventions are depicted in
Fig.~\ref{fig:affine_param}.

\subsection{Neutrinosurface Approximation}
\label{ssec:neutrinosurface}
\todo{derive $\tau=2/3$ approx \cite[Chap.~4]{shu1991-book_radiation},
  and \cite{miln1921-radiative_equil_1}}
In the simplified configuration of a plane-parallel atmosphere, the
rendering equation reduces to 
\begin{equation}
\label{eqn:neutrinosurface}
  f(\lambda) = f^{\rm eq}(\lambda_0),
\end{equation}
where $\tau(\lambda,\lambda_0)=2/3$.

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{clarify integral differentials}
We may take angular moments of the distribution function:
\begin{align}
  J(\varepsilon) &=
  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  H^\mu(\varepsilon) &=
  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
  S^{\mu\gamma}(\varepsilon) &=
  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
\end{align}
which represent energy density, energy flux, and momentum flux, respectively.
\todo{clarify moment frame and relate to other frames}

\subsection{Numerical Implementation}
\label{ssec:numerical}
Much of our numerical implementation is borrowed from the geodesic evolution
system described in~\cite{bohn2016-code}.
\todo{describe eqns for $f_{\rm integ}$ and $\tau$}
\todo{describe $\nu$-specific numerics}

@@

\subsection{Dummy Tests}
Maybe we don't need to present these tests, but make sure to establish
confidence in the basics:
gravitational redshift, Doppler shift, thermodynamic equilibrium.
\todo{setup and run dummy tests}

\subsection{Homogenous Absorbing Star}
Test described in~\cite[Sec.~3.2]{smit1997-two_moment}
\todo{Jerred: run test}

\subsection{Homogenous Scattering Star}
Test presented in~\cite{humm1971-grey_transfer}
and described in~\cite[Sec.~9.1.1]{abdi2012-monte_carlo}.
\todo{formulate analytic $J$ and $F^\mu$}
\todo{Jerred: setup and run test}

\subsection{Post-Bounce Neutron Core}
Standard radiation transport test described in~\cite{ocon2015-gr1d_with_nu},
and in \cite[App.~E.6]{fouc2015-m1_nsbh},
and in~\cite{abdi2012-monte_carlo}.
\todo{Evan: provide initial data and $H^r(\varepsilon)$}
\todo{Jerred: setup and run test}

\subsection{Disk Comparison to M1}
Compute total and relative number and energy luminosities for a disk evolved
in one of~\cite{fouc2015-m1_nsbh, fouc2016-m1_nsns, fouc2016-m1_evolve_n}.
Expect agreement within factors of a few (?).
\todo{code up moment integrals}
\todo{compute luminosities from M1}
\todo{compute luminosities from ray tracing}

@@

\subsection{Emission Spectra}
\label{ssec:spectra}
The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
Something like this is presented in~\cite[Fig.~10-11]{pere2014-nu_wind};
compare.
\todo{code up moment integrals}
\todo{compute $H^r(\theta;\varepsilon)$ in nsns and nsbh}
\todo{maybe compute at several times}

\subsection{Closure Relation}
\label{ssec:closure}
\todo{figure out your approach}
This is a little tricky. Unless we can come up with a new improved analytic
closure relation and use ray tracing to evaluate its trustworthiness,
it's going to be hard to get anything useful from computing
$S^{\alpha\beta}$.

Maybe the variable Eddington closure is more suitable to this application.
If I understand it right, applying ray tracing to that method would require
computing $S^{\alpha\beta}$ on the fly throughout the spatial volume.
That would certainly be doable for a quasi-stationary disk configuration.
\todo{compute $S^{\alpha\beta}$ and evaluate its effect on evolution}

@@

\section{Source Terms}
\label{sec:source_terms}
\todo{derive and type up sources in terms of $J$ and $H^\alpha$
  following \cite[Sec.~4]{shib2011-truncated_moment}}
Here we present the sources comprising the right hand side of the
Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) in terms of the microphysical
processes of absorption/emission, elastic scattering, inellastic scattering,
and pair annihilation/production
(detailed in Tab.~\ref{tab:neutrino_processes}).
